---
title: "Project1.1"
author: "Shubhi Saxena"
date: "12/4/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
library(lubridate)
```

# Reading the files

```{r}
dir1 <- "D:/IDMP/Project"
f_data <- read_csv(file.path(dir1 , "features.csv"))
s_data <- read_csv(file.path(dir1 , "stores.csv"))
train_data <- read_csv(file.path(dir1 , "train.csv"))
test_data <- read_csv(file.path(dir1, "test.csv"))
in_data <- inner_join(f_data , s_data , by = c("Store"))
wal_data <- inner_join(in_data , train_data , by = c("Store", "Date" , "IsHoliday"))
```

# Preprocessing the Data

-> Changing th Holiday Variable

```{r}
wal_data$IsHoliday[wal_data$IsHoliday=="TRUE"]<-1
wal_data$IsHoliday[wal_data$IsHoliday== "FALSE"]<-0
head(wal_data)
```
-> Check for NAs
```{r}
sum(is.na(wal_data))
```
```{r}
colSums(is.na(wal_data))
```
-> Dealing with NAs
```{r}
wal_data[is.na(wal_data)] <- 0 
```
```{r}
colSums(is.na(wal_data))
```
-> Keeping only positive weekly sales data

```{r}
wal_data<- wal_data[wal_data$Weekly_Sales> 0, ]
```

# Univariate Analysis

-> Markdown Variable analysis

```{r}
Mr1<-aggregate(MarkDown1 ~ Date , data=wal_data, sum)
Mr2<-aggregate(MarkDown2 ~ Date , data=wal_data, sum)
Mr3<-aggregate(MarkDown3 ~ Date , data=wal_data, sum)
Mr4<-aggregate(MarkDown4 ~ Date , data=wal_data, sum)
Mr5<-aggregate(MarkDown5 ~ Date , data=wal_data, sum)
```

```{r}
#png("MarkdownTrend.png", width = 800, height = 550, res=150)
col=c("MR1"="red","MR2"="blue","MR3"="green","MR4"="yellow","MR5"= "orange")
g1<- ggplot()+
  geom_line(data=Mr1,aes(x=Date,y=MarkDown1,color="MR1"),size=0.8)+
  geom_line(data=Mr2,aes(x=Date,y=MarkDown2,color='MR2'),size=0.8)+
  geom_line(data=Mr3,aes(x=Date,y=MarkDown3,color='MR3'),size=0.8)+
  geom_line(data=Mr4,aes(x=Date,y=MarkDown4,color='MR4'),size=0.8)+
  geom_line(data=Mr5,aes(x=Date,y=MarkDown5,color='MR5'),size=0.8)+
  scale_x_date(breaks= seq(as.Date("2010-02-05"),
                           as.Date("2013-01-01"),by="4 months"),
               date_labels="%b\n%Y")+
  labs(title="Trend of Markdown all over the dataset", x="Date",
       y="value of markdown ")
  g1+ scale_color_manual(values=col)+theme_minimal()
  


```
Inference:

1. We can see that till December 2011 there were no markdown values available

2.All markdown values are giving a spike at start like in month of january 2012 and in start markdown 3 is highest but after start markdown 1 is high throughout dataset

-> Type of stores

```{r}
typeA <- wal_data[wal_data$Type=='A',]
type1<- aggregate(Weekly_Sales ~Date,data=typeA, sum)
typeB <- wal_data[wal_data$Type=='B',]
type2<- aggregate(Weekly_Sales ~Date,data=typeB, sum)
typeC <- wal_data[wal_data$Type=='C',]
type3<- aggregate(Weekly_Sales ~Date,data=typeC, sum)

```

```{r}
vals <- c("typeA" ="red","typeB"="blue","typeC"="green")
gg<- ggplot()+
  geom_line(data=type1,aes(x=Date,y=Weekly_Sales,color="typeA"),size=0.8)+
  geom_line(data=type2,aes(x=Date,y=Weekly_Sales,color="typeB"),size=0.8)+
  geom_line(data=type3,aes(x=Date,y=Weekly_Sales,color="typeC"),size=0.8)+
  scale_x_date(breaks= seq(as.Date("2010-02-05"),
                           as.Date("2013-01-01"),by="6 months"),
               date_labels="%b\n%Y")+
  labs(title="Trend of type of Store and weekly sales",
       x = "Date",y="Weekly Sales")
 

gg+ scale_color_manual(values = vals)+theme_minimal()

  
```
Inference:

1. There is sesonality in weekly sales of Type A and Type B with increase in weekly sales around november and december of each year

2.Type C store weekly sales doesn't rise just like type A and type B ,it is fluctuating around its mean values only

-> Temperature

```{r}
temp<- aggregate(Temperature ~Date,data=wal_data, mean)
gg<- ggplot()+
  geom_line(data=temp,aes(x=Date,y=Temperature),size=0.8)+
  scale_x_date(breaks= seq(as.Date("2010-02-05"),
                           as.Date("2013-01-01"),
                           by="6 months"),
               date_labels="%b\n%Y")+theme_minimal()
  
gg
```
# Multivariate Analysis

1. Sales and Store Number
```{r}
#png("Stores.png", width = 800, height = 550, res=150)
wal_data %>%
  ggplot(aes(x=as.factor(Store),Weekly_Sales, fill=as.factor(Store)))+
  geom_boxplot()+
  ylim(c(20000,80000))+
  labs(title="Box Plot of Weekly Sales by Store Number",
       x="Store",y="Weekly Sales in Thousands")+
  theme(axis.text.x = element_text(angle = 0, vjust = 0.8, 
    size = 10),legend.position = "none")
#dev.off()
```
Inference:

1.We can see that store 4,10 and 20 has highest weekly sales throughout and average weekly sales is around 10000

2.We can plot type als0 with store to see what is the type of store which are creating big impact on weekly sales


2. Sales, Store Number and its type
```{r}
wal_data %>%
  ggplot(aes(x=as.factor(Store),Weekly_Sales, fill=as.factor(Type)))+
  geom_boxplot()+
  ylim(c(20000,80000))+
  labs(title="Box Plot of Weekly Sales by Store Number and its Type",x="Store",y="Weekly Sales in Thousands",
       fill="Type")+
  theme(axis.text.x = element_text(angle = 0, vjust = 0.8, 
    size = 10),legend.position = "top")
```
Inference:

1.We can interpret from box plot that mean of most of the type A store are higher as compared to type B and C .

2.Therefore we can say that type A store will be dominant in overall weekly sales from each store


3. Sales, Store Number and IsHoliday

```{r}
wal_data %>%
  ggplot(aes(x=as.factor(Store),Weekly_Sales, fill=factor(IsHoliday,label=c("False","True"))))+
  geom_boxplot(outlier.shape = NA)+
  ylim(c(0,100000))+
  labs(title="Box Plot of Weekly Sales by Store Number and Holiday",x="Store",
       y="Weekly Sales in Thousands",
       fill="IsHoliday")+
  theme(axis.text.x = element_text(angle = 0, vjust = 0.8, 
    size = 10),legend.position = "top")
```
Inference:

1.From plot we can see that weekly sales of all store on holiday is higher as compared to non holiday week.

2.So,yes holiday weeks are creating huge impact and can be treated as important feature also in determining the weekly sales

